---
title: "**Menstrual Health & Sexual Education In India**"
author: "_Group 2: Kriti Vajjhula, Franchezca Layog, Ian Chandramouli, & Harmeet Singh_"
date: "_March 2021_"
---

```{r setup, include = FALSE}

library('knitr')
source("analysis.R")

```

## **Problem Domain Description**

Millions of women in India face great challenges to managing their menstrual 
health and practicing menstrual hygiene. This is due to a number of enabling 
factors present in the country's distribution system for products, income 
level, and social stigma. One of the largest factors is the lack of education 
and awareness to young women in the process of menstruation and its physiological 
basis pre-menarche (before a young girl's first menstrual cycle). Another key
issue is the availability of products to all income levels. The pricing of 
commercial products are expensive as well as inaccessible to women in rural areas,
who account for a significant portion of the country's population. Sanitation
routines, attitudes, and most importantly, facilities, are also very underdeveloped
across the country and limit access to functioning toilets and disposal solutions. 
This harms working women as well as school going young women, greatly increasing 
the gender income gap in income as well as academia.

Our group is interested in delving into the sociocultural root of this problem 
and finding correlations, insights, and potential solutions for dismantling 
these enabling factors. We will attempt to focus on the cause of these enabling 
factors and finding connections in data between socioeconomic backgrounds, 
income levels, religion, caste, etc.

The first existing project we found is a country landscape analysis of menstrual health 
in India. The report seeks to understand the current state of womens' health 
in India and includes research from interviews from adolescent girls from rural 
and urban areas.

Source: [Menstrual Health in India - Country Landscape Analysis](https://www.susana.org/en/knowledge-hub/resources-and-publications/library/details/2581)

The second existing project we examined studies the effect of a community-based health 
education intervention on change in behavior, routines, and awareness in the menstrual 
hygiene of rural adolescent girls.

Source: [The Effect of Community-Based Health Education Intervention on Management of Menstrual Hygiene among Rural Indian Adolescent Girls](https://www.longwoods.com/content/19303/world-health-population/the-effect-of-community-based-health-education-intervention-on-management-of-menstrual-hygiene-among)

## **Data Description**

### Percentage of Women Complication During Pregnancy Data Set
Source: [Percentage of Women Complication During Pregnancy](https://data.gov.in/resources/percentage-women-who-had-any-pregnancy-complication-any-delivery-complication-any-post)

This data set describes the percentage of women who had pregnancy complications 
according to different reasons such as menstrual related problems or delivery complications.
The data was obtained by the Ministry of Health and Family Welfare of India. 
The data was collected via a nationwide survey in which Women answered a variety of questions.
This data would hopefully answer the question as to which policies are effective 
in women's menstrual healthcare. This dataset provides information about 
individual districts which can be used jointly with another dataset to answer 
our questions.

The data set will look like this:

```{r preg_comp_df_sample, echo = FALSE}

kable(head(preg_comp_df))

```

*The above table is a sample of the data set preg_comp_df.*

The column names for this data set: <br> 
- preg_comp = % Of Women with Any Pregnancy Complications <br>
- any_delivery_comp = % of Woman with Delivery Complications <br>
- any_post_delivery_comp = % of Women with Post-Delivery Complications <br>
- vag_discharge_comp = % of Women with Problems with Vaginal Discharge in Last 3 Months <br>
- menstrual_related_comp = % of Women with Menstrual Related Problems in Last 3 Months <br>

#### Summary Statistics (preg_comp_df)


``` {r, echo = FALSE}
mens_comp_state_plot
```

*The above plot describes the distribution of menstrual related complications 
during pregnancy cases across various states from the pre_comp_df data frame.*

#### Outliers

A notable outlier is `r mens_comp_state_df %>% filter(menstrual_related_comp == max(menstrual_related_comp)) %>% pull(States)`. This particular value is notable due to the fact that it indicates an unusually large amount of menstrual complications in relation to the rest of the data set. This can be attributable to the fact that West Bengal is one of the most populous states in the country. 

### Menstrual Hygiene Scheme Budget 2016-2020 Data Set
Source: [Menstrual Hygiene Scheme Budget](https://data.gov.in/resources/stateut-wise-menstrual-hygiene-scheme-budget-approved-under-nhm-2016-17-2019-20-ministry)

The data was released by Rajya Sabha (Council of States), the upper house of the 
Indian Parliament. The government is logically the entity responsible for proposing 
the budget, and the data was therefore effectively generated by them. It was 
released under the National Data Sharing and Accessibility Policy.
This will help us identify the financial disparities in government menstrual
health support programs by state, which can provide insight as to how attitudes 
toward and the overall state of female menstrual health differ by state and region.

The data set will look like this:

```{r menstrual_budget_df_sample, echo = FALSE}

kable(head(menstrual_budget_df))

```

*The above table is a sample of the data set menstual_budget_df*

The column names for this data set: <br>
- state_no = State No. in Alphabetical Order <br>
- 2016-17 Budget <br>
- 2017-18 Budget <br>
- 2018-19 Budget <br>
- 2019-20 Budget <br>

#### Summary Statistics (total_budgets_df)

```{r total_budgets_plot, echo = FALSE}

total_budgets_plot

```

*The above plot visualizes the average budget spent by all the states in the 
data set on menstrual related policies*

#### *2016-2017*

The average spending for the years of 2016-2017 is `r mean(mens_budget_2016_17)`.

The range of this data set is `r max(mens_budget_2016_17) - min(mens_budget_2016_17)`.

This data set comes with `r mens_budget_2016_17['NA\'s']` NA values.

#### *2017-2018*

The average spending for the years of 2017-2018 is `r mean(mens_budget_2017_18)`.

The range of this data set is `r max(mens_budget_2017_18) - min(mens_budget_2017_18)`.

This data set comes with `r mens_budget_2017_18['NA\'s']` NA values.

#### *2018-2019*

The average spending for the years of 2018-2019 is `r mean(mens_budget_2018_19)`.

The range of this data set is `r max(mens_budget_2018_19) - min(mens_budget_2018_19)`.

This data set comes with `r mens_budget_2018_19['NA\'s']` NA values.

#### *2019-2020*

The average spending for the years of 2019-2020 is `r mean(mens_budget_2019_20)`.

The range of this data set is `r max(mens_budget_2019_20) - min(mens_budget_2019_20)`.

This data set comes with `r mens_budget_2019_20['NA\'s']` NA values.

#### Outliers

For each fiscal year, there was a lot of missing data. At the maximum, there was data
present for 17 states.

Overall, the data seemed to be fairly consistent for what was present however, there is an
evident outlier for the year of 2019-20 as the average is significantly higher than the rest.
If we take a closer look at our data set, we can see that there is an entry for Rajasthan
that was higher than others provided. Additionally, it was also the only year Rajasthan
provided any data indicating that there was an issue with data collection somewhere along the line.
This same phenomena arose a couple of times throughout the data set.

All NAs were replaced with 0s for calculations which could potentially misrepresent a budget
that was actually there but not reported. In some cases, the could drive the average either
up or down.

``` {r mens_budget_line_plot, echo = FALSE}
suppressMessages(plot(mens_budget_line_plot))
```

*The above plot visualizes the budget for each state (with no missing values in the
the data set) from 2016-2020.*

#### Outliers

We decided to use this form of plot to show the distribution of data to show the progression of 
budget changes throughout the years of data collection by the states which have the most complete
data values without any NA values. An outlier is shown within the budget line for Maharashtra, there is a steep decline for the data point of the years of 2018-19. The other states' budget point progressions are relatively more stable and don't have sudden budget changes. 

## Specific Question Analyses

### States Spending vs. Menstrual Related Pregnancy Complications

**How does  the amount of each state's budget for a particular year impact the percentage amount of menstrual related complications?**

- Introduction:
  - For this question, our group wanted to investigate the different budget amounts spent by each state in the 2016-17 time frame to see if they had any correlation to the number of menstrual related pregnancy complications. We wanted to see if particular states were investing more efforts monetarily towards menstrual health, if that meant there would be less occurrences of menstrual related pregnancy complications.
  
- Data Wrangling:
  - Firstly, to begin our data wrangling, we had to manipulate our pregnancy complications data frame to filter out for the totals of each specific
  complication from each state. We then created a new data set by joining the pregnancy totals data frame with our menstrual budget data frame,
  joining by states and filtered by state, budget, and the total menstrual budget complications. <br>

```{r preg_budget_df, echo = FALSE}

kable(preg_budget_df)

```

```{r preg_budget_plot, echo = FALSE}

preg_budget_plot

```

- Results Evaluation: 
  - The data in the graph is inconclusive in answering the analysis question. 
  - Given our data sets, the joined data is not an accurate representation of the 
population of India. There are only 11 out of the 23 states represented. This may be to the lack of responses of budget data and pregnancy complication data from particular states. 
  - Additionally, there are some obvious outliers in the data seen in the graph further down the x-axis. West Bengal and Maharashtra are obvious outliers because they don't seem to be following the trend of decreasing number of complications with increasing budget. 
  - However, currently if a line of best fit were to be drawn, there would be a loose 
correlation between the budget spent in 2016-17 and the number of menstrual 
related pregnancy complications present. It is yet to be determines if these are 
truly correlated or if there may be a hidden variable such as birth rate or access 
to healthcare. 
  - In the end, we can not confidently answer our analysis question without more sufficient
data from a wider sample of India's population. 


### States Spending vs. Total Percentage Pregnancy Complications

**How does the does each states' budget correlate to the percentage of overall pregnancy complications?**

- Introduction: 
  - For this question, our group wanted to investigate the different 
budget amounts spent by each state in the 2016-17 time frame to see if they had
any correlation to the total percentage amount of pregnancy complications. We used
the same percentage of women with pregnancy complications data set as the last analysis
question. Similar to the last analysis question, we wanted to be able to compare if menstrual
related pregnancy complications play a large role in the total percentage amount
of pregnancy complications. 

- Data Wrangling: 
  - Firstly, to begin our data wrangling, we had to manipulate our pregnancy 
  complications data frame to filter out for the totals of each specific
  complication from each state. We then created a new data set by joining the 
  pregnancy totals data frame with our menstrual budget data frame,
  joining by states and filtered by state, budget, and the total percentage 
  of pregnancy complciations <br>
  
```{r preg_comp_budget_plot, echo = FALSE}

preg_comp_budget_plot

```

- Results Evaluation:
  - Similar to the last graph, we cannot make any definite assumptions based on the
  data set. The current data is not representative of India because out of the 28
  total states, 23 were present in the data frame and only 11 were  present in both 
  data sets when joined.
  - Additionally, there are a few outliers in the data, explicitly seen in the graph. 
  The states Tripura, West Bengal, Haryana, Maharashtra do not seem to be following the
  trend of decreasing compliations by increasing budget.
  - Regardless of the outliers, there seems to be loose correlation for the rest
  of the states. This may be due to unforseen factors such as the ones stated above in
  the previous analysis question.
  - In conclusion, we cannot make any assumptions that menstrual-related pregnancy 
  complications make any significant impact on the overall percentage of pregnancy
  complications without more sufficient data from a wider range of data from all of India.

### Distribution of State Menstrual Health Budget in 2019-2020

**What does the distribution of the menstrual health budget look like in 2019-20 across different states in the data set?**

- Introduction:
  - Our group created this analysis question to investigate the most recent menstrual budget 
  data and how the quantities of spending are broken down by state. We wanted to be able to get
  an idea of the spending by state and the comparison between budgets for the most recent time 
  frame. We figured that the best way to visualize this is using a pie chart in order to
  understand the size of the budget distribution with different slices of the pie. 

- Data Wrangling:
  - For our data wrangling, we first filtered menstrual health budget data frame for
the year 2019-20 where the budget is greater than 0. We then created a pie chart
that would easily visualize the distribution of each state's budget from the year
in relation to each other. <br>

```{r budget_2019_20_df, echo = FALSE}

kable(budget_2019_20)

```

```{r budget_2019_20_plot, echo = FALSE}

budget_2019_20_plot

```

- Results Evaluation: 
  - The data in the pie chart is inconclusive in answering our data analysis question.
  - The most prominent outlier is the state of Rajasthan because they provided the largest
spending budget, but it was also the state's only entry in the data frame. 
  - Another stipulation is that the data set is the graph is not accurately representative
of the entire population of India. There is ambiguity as to whether some regions listed
in the data set are states versus cities. Additionally, there is the exclusion of 
some states. 
  - In the end, we can not confidently answer our analysis question without more sufficient
data from a wider sample of India's population. 


### Distribution of Pregnancy Complications by State

**What is the distribution of pregnancy complications by state?**

- Introduction:
  - As a group, wanted to be able to pinpoint the most common types of pregnancy complication by
  state. Therefore, we decided to visualize the percentages of complications across the 
  data time period as a faceted bar graph. The distribution of pregnancy complications by state 
  is easily comparable in this manner and summarizes the most common type of complication in an 
  easy-to-read fashion. 

- Data Wrangling:
  - For our data wrangling, we decided to group by state, use a summarize function to aggregate all of the values for each type of complication and created a new column to include all of these values. Finally, we used the pivot_longer() function to include percentages for each complication. 

```{r preg_comp_by_state_df, echo = FALSE}
kable(head(preg_comp_by_state_df))
```

```{r preg_comp_bar_graph, echo=FALSE}
preg_comp_bar_graph
```

- Results Evaluation:
  - One of the first things to note when evaluating the results is the surveying 
  method that was most likely used. When adding up the various pregnancy complications they exceed 
  100%. This leads us to conclude that the complications are not mutually exclusive, implying that
  a woman could have post-delivery complications while also experiencing menstrual-related
  complications. 
  - There is a positive correlation between the length of the bars and the most populous states in
  India. This means that there are more numbers of complications of all types within states that have
  higher populations, such as West Bengal, Tamil Nadu, and Karnataka. 
  - An important thing to note is that the results displayed by our plot do not show a fully accurate
  picture of the real number of complications across India, rather an aggregation of percentages. 
  
  
