---
title: "**Menstrual Health & Sexual Education In India**"
author: "_Group 2: Kriti Vajjhula, Franchezca Layog, Ian Chandramouli, & Harmeet Singh_"
date: "_March 2021_"
---

```{r setup, include = FALSE}

source("analysis.R")

```

## Problem Domain Description

Millions of women in India face great challenges to managing their menstrual 
health and practicing menstrual hygiene. This is due to a number of enabling 
factors present in the country's distribution system for products, income 
level, and social stigma. One of the largest factors is the lack of education 
and awareness to young women in the process of menstruation and its physiological 
basis pre-menarche (before a young girl's first menstrual cycle). Another key
issue is the availability of products to all income levels. The pricing of 
commercial products are expensive as well as inaccessible to women in rural areas,
who account for a significant portion of the country's population. Sanitation
routines, attitudes, and most importantly, facilities, are also very underdeveloped
across the country and limit access to functioning toilets and disposal solutions. 
This harms working women as well as school going young women, greatly increasing 
the gender income gap in income as well as academia.

Our group is interested in delving into the sociocultural root of this problem 
and finding correlations, insights, and potential solutions for dismantling 
these enabling factors. We will attempt to focus on the cause of these enabling 
factors and finding connections in data between socioeconomic backgrounds, 
income levels, religion, caste, etc.

# Data Description

## Percentage of Women Complication During Pregnancy Data Set
Source: [Percentage of Women Complication During Pregnancy](https://data.gov.in/resources/percentage-women-who-had-any-pregnancy-complication-any-delivery-complication-any-post)

This data set describes the percentage of women who had pregnancy complications 
according to different reasons such as menstrual related problems or delivery complications.
The data was obtained by the Ministry of Health and Family Welfare of India. 
The data was collected via a nationwide survey in which Women answered a variety of questions.
This data would hopefully answer the question as to which policies are effective 
in women's menstrual healthcare. This dataset provides information about 
individual districts which can be used jointly with another dataset to answer 
our questions.

The data set will look like this:

```{r preg_comp_df_sample, echo = FALSE}

head(preg_comp_df)

```

The column names for this data set: <br> 
- any_preg_comp = % Of Women with Any Pregnancy Complications <br>
- any_delivery_comp = % of Woman with Delivery Complications <br>
- any_post_delivery_comp = % of Women with Post-Delivery Complications <br>
- vag_discharge_comp = % of Women with Problems with Vaginal Discharge in Last 3 Months <br>
- menstrual_related_comp = % of Women with Menstrual Related Problems in Last 3 Months <br>

## Menstrual Hygiene Scheme Budget 2016-2020 Data Set
Source: [Menstrual Hygiene Scheme Budget](https://data.gov.in/resources/stateut-wise-menstrual-hygiene-scheme-budget-approved-under-nhm-2016-17-2019-20-ministry)

The data was released by Rajya Sabha (Council of States), the upper house of the 
Indian Parliament. The government is logically the entity responsible for proposing 
the budget, and the data was therefore effectively generated by them. It was 
released under the National Data Sharing and Accessibility Policy.
This will help us identify the financial disparities in government menstrual
health support programs by state, which can provide insight as to how attitudes 
toward and the overall state of female menstrual health differ by state and region.

```{r menstrual_budget_df_sample, echo = FALSE}

head(menstrual_budget_df)

```

The column names for this data set: <br>
- state_no = State No. in Alphabetical Order <br>
- 2016-17 Budget <br>
- 2017-18 Budget <br>
- 2018-19 Budget <br>
- 2019-20 Budget <br>

## Specific Question Analyses